Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-14 paper/paper.tex
pass

\subsubsection{Frequency of content modifiers.} Content options are
associated with specific packets fields\Mar{or modifiers?}.\Mar{why
  GET is associated with http\_method and http\_header?}\Mar{are all
  content associated with one or many fields or positions in a packet?
  can you give examples other than GET to explain this? if yes,
  shouldn't this be discussed before? Maybe Section 2.3?}\Mar{Also, it
  is important to justify (to explain rationale)---how is it different
  from counting rarity?}

@@

\noindent
, where $N$ is the number of contents in a given rule and $M$
is the number of content modifiers\Mar{in a rule? what is a content
  modifier in a rule?}.

\begin{figure}[H]
  \centering
  \scalebox{1.2}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-2ex}
  \caption{\label{fig:distribution-content_modifiers}Content modifiers
    frequency.\Mar{This is awkward. This is the only histogram that
      only shows 6 values in the x-axis. Aren't there more?}}
\end{figure}

>>>>

\subsubsection{Frequency of content modifiers.} As previously
mentioned, \percRulesWithContent\ of Suricata rules involve content
options. Furthermnore, we observed that many of these rules indicate
what parts of the message the content string should be checked
against.\Mar{Lucas, (1) vc. tem dados para dar suporte a isto (de http
  e tcp)? ou seja, quantas options do tipo ``content'' possuem
  modificadores.}  For example, the creator of the rule may want to
indicate that a specific string be present in the header of an HTTP
request message. Content \emph{modifiers} serve this purpose. For
example, the modifier \CodeIn{http\_header} in the option
\CodeIn{content: ``keep-alive''; http\_header} states the requirement that
the string ``keep-alive'' be present in the header of the message. (A
keep-alive connection attribute indicates that a single TCP connection
should remain open for multiple HTTP requests/responses.)
Figure~\Fix{adicione pequena figura...verifique se ainda faz sentido
  depois do que falei acima...}

Suricata supports content modifiers for tcp and http. Given that
content options often use modifiers\Mar{assumindo que eh
  verdade. confirmar estat. solicitada acima...}, \tname\ uses
associated frequency to discriminate more or less likely rules. For
example, historical data indicates that content options that use the
\CodeIn{http\_connection} modifiers are much less likely than content
options that use the modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the distribution of content modifiers for http. We ommit the
distribution for tcp for space. Similarly to the the previous
heuristic functions, this function uses the frequency of content
modifiers to calibrate the contribution of each rule.

\begin{figure}[H]
  \centering
  \scalebox{1.2}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-2ex}
  \caption{\label{fig:distribution-content_modifiers}Content modifiers
    frequency.\Mar{what about ``none'' (no modifier)?}}
\end{figure}

%% \section{Modificadores}É possivel deixar que uma regra verifique todo o payload do pacote em busca de um match com um content ou especificar partes do payload onde esse content deve ser procurado. Essa definição é feita utilizando os chamados modificadores de content, eles adicionam mais detalhes/informações sobre a detecção de um determinado content. Os modificadores que estamos considerando são validos apenas para regras dos protocolos tcp e http.

%% Exemplo do payload "cru" de um pacote http:\\
%% Aqui, todas as informações do pacote estão juntas.

%% \begin{figure}[h!]
%%   \lstinputlisting[language=C,numbers=none,keywords={content}]{http-example.payload}
%% \end{figure}

%% O mesmo payload interpretado pelo Wireshark:\\
%% Aqui, o wireshark interpreta as informações dos pacotes e as divide para melhor visualização e entendimento.

%% \begin{figure}[H]
%%   \lstinputlisting[language=C,numbers=none,keywords={content}]{http-example-wireshark.payload}
%% \end{figure}

%% O que o suricata faz eh identificar os campos e seus valores e com isso permite que as regras façam referencia direta a eles. Por exemplo, na figura do payload "cru" não conseguimos identificar facilmente oq eh o "Request URI Path:", já na segunda figura isso está explícito.

%% Eu disse que o "GET" estava associado tambem ao http\_header, mas não está, me enganei. Mas sim, pode haver interseção entre os campos que dois ou mais modificadores referenciam. O seguinte trecho do payload é acessado pelo modificador http\_header:

%% \begin{figure}[H]
%%   \lstinputlisting[language=C,numbers=none,keywords={content}]{http-header.payload}
%% \end{figure}

%% Entretando, tambem existem os modificadores http\_host,
%% http\_accept\_enc, http\_connection e http\_accept que permitem
%% acessar diretamente cada uma das linhas acima.

\Mar{----------------------------------------------------------- estou
aqui}

\subsection{Pontuação}
A avaliação que fazemos aqui pontua um dado content de acordo com quantas vezes o campo em que ele se encontra eh referenciado nas regras. Por exemplo, vamos supor que o tokenizer produziu a opção content: "/HtmlAdaptor". Essa string pertence ao campo "Request URI:". Esse campo pode ser acessado pelo modificador http\_uri. O tokenizer associa o content a esse modificador. Posteriormente, esse content vai receber uma pontuação correspondente a popularidade desse modificador:\\
\[\frac{10111}{10111*1}\]
A multiplicação por 1 no denominador eh referente a quantidade de modificadores que estão associadas a esse content.
No caso da opção content: "keep-alive", essa string pertence ao campo "Connection:". Como falado acima, o suricata pode acessar esse campo pelos modificadores http\_header ou http\_connection. Portanto, esse content será pontuado de acordo com a popularidade desses dois modificadores:
\[\frac{4089+29}{10111*2}\]

Content options are
associated with specific packets fields\Mar{or modifiers?}.\Mar{why
  GET is associated with http\_method and http\_header?}\Mar{are all
  content associated with one or many fields or positions in a packet?
  can you give examples other than GET to explain this? if yes,
  shouldn't this be discussed before? Maybe Section 2.3?}\Mar{Also, it
  is important to justify (to explain rationale)---how is it different
  from counting rarity?}

@@

\noindent
, where $N$ is the number of contents in a given rule and $M$
is the number of content modifiers\Mar{in a rule? what is a content
  modifier in a rule?}. \Luc{M eh o numero de modificadores associados
  a cada content. Temos content[1], content[2], ..., content[N]. Cada
  um desses contents tem modificadores associados content[i][1],
  content[i][2], ..., content[i][M].}
